---
title: "Project: EHR SOFA --- prediction of mortality among ventilated"
subtitle: "Explore SOFA values that were imputed to zero"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: "H:\\word rMarkDown template.docx"
---

```{r global_options, include=FALSE}
library(partykit)
library(dplyr)
library(pROC)
library(ROCR)
library(caret)
library(RColorBrewer)
library(flextable)
library(rmarkdown)
library(knitr)
library(readr)

library(givitiR)
library(gridExtra)
library(ggplot2)

library(gtsummary)
library(flextable)
library(officer)

opts_chunk$set(fig.width=10, fig.height=8,echo=FALSE, warning=FALSE, message=FALSE, comment = NA)
```

```{r package_options, include=FALSE}
opts_knit$set(progress = TRUE, verbose = TRUE)
```


```{r}
load(here::here("FirstMV_20JUL2021_final.RData"))
```

## Conditional Tree

```{r}
### remove patients with SOFA score imputed as 0
FirstMV.select <- FirstMV.select%>%filter(!is.na(Pre1day_total_score_num))
FirstMVdata.train<- FirstMVdata.train%>%filter(!is.na(Pre1day_total_score_num))
FirstMVdata.test <- FirstMVdata.test%>%filter(!is.na(Pre1day_total_score_num))

FirstMV.select <- FirstMV.select%>%
  mutate_if(is.character, as.factor)%>%
  mutate(FourGrp_Pre1day_total_score_num = case_when(
    is.na(Pre1day_total_score_num) ~ NA_character_,
    Pre1day_total_score_num<6 ~ "sofa <6",
    Pre1day_total_score_num>=6 & Pre1day_total_score_num<9 ~ "6 >= sofa <9",
    Pre1day_total_score_num>=9 & Pre1day_total_score_num<12 ~ "9 >= sofa <12",
    Pre1day_total_score_num>=12 ~ "sofa >= 12"
  ),
  FourGrp_Pre1day_total_score_num = factor(FourGrp_Pre1day_total_score_num, levels=c("sofa <6","6 >= sofa <9","9 >= sofa <12","sofa >= 12")))%>%
  select(-contains("_imp"))
 

FirstMVdata.train <- FirstMVdata.train%>%
  mutate_if(is.character, as.factor)%>%
  mutate(FourGrp_Pre1day_total_score_num = case_when(
    is.na(Pre1day_total_score_num) ~ NA_character_,
    Pre1day_total_score_num<6 ~ "sofa <6",
    Pre1day_total_score_num>=6 & Pre1day_total_score_num<9 ~ "6 >= sofa <9",
    Pre1day_total_score_num>=9 & Pre1day_total_score_num<12 ~ "9 >= sofa <12",
    Pre1day_total_score_num>=12 ~ "sofa >= 12"
  ),
  FourGrp_Pre1day_total_score_num = factor(FourGrp_Pre1day_total_score_num, levels=c("sofa <6","6 >= sofa <9","9 >= sofa <12","sofa >= 12")))%>%
  select(-contains("_imp"))
  
  
FirstMVdata.test <- FirstMVdata.test%>%
  mutate_if(is.character, as.factor)%>%
  mutate(FourGrp_Pre1day_total_score_num = case_when(
    is.na(Pre1day_total_score_num) ~ NA_character_,
    Pre1day_total_score_num<6 ~ "sofa <6",
    Pre1day_total_score_num>=6 & Pre1day_total_score_num<9 ~ "6 >= sofa <9",
    Pre1day_total_score_num>=9 & Pre1day_total_score_num<12 ~ "9 >= sofa <12",
    Pre1day_total_score_num>=12 ~ "sofa >= 12"
  ),
  FourGrp_Pre1day_total_score_num = factor(FourGrp_Pre1day_total_score_num, levels=c("sofa <6","6 >= sofa <9","9 >= sofa <12","sofa >= 12")))%>%
  select(-contains("_imp"))



myTreeFunc <- function(var,...){
  
set.seed(123)
temp0 <- (ctree(death~.,data=FirstMV.select%>%select(...),minbucket=1000))

plot(temp0, tp_args = list(fill =c("darkgreen","red")))

# Evaluate model performance On training set
prob.discharge <- predict(temp0, newdata=FirstMVdata.train, type="prob",simplify=T)[,2]
roc.train <- roc(FirstMVdata.train$death, prob.discharge)
ci.train <- ci.auc(roc.train)
auc.train <- paste0(round(as.numeric(roc.train$auc),3), " (",round(as.numeric(ci.train)[1],3),",",round(as.numeric(ci.train)[3],3),")")

# evaluate model performance On test set
prob.discharge <- predict(temp0, newdata=FirstMVdata.test, type="prob",simplify=T)[,2]
roc.test <- roc(FirstMVdata.test$death, prob.discharge)
ci.test<- ci.auc(roc.test)
auc.test <- paste0(round(as.numeric(roc.test$auc),3), " (",round(as.numeric(ci.test)[1],3),",",round(as.numeric(ci.test)[3],3),")")

# confusion matrix 
# predict.discharge <- predict(temp0, newdata=FirstMVdata.test, type="response",simplify=T)
# CM <-confusionMatrix(as.factor(FirstMVdata.test$death),predict.discharge)
# CM$table
out <- data.frame(Variable = var,AUC.train = auc.train, AUC.test = auc.test)
return(kable(out))
}
```



### SOFA + all components

```{r}
myTreeFunc(var = "SOFA + all components" ,contains("Pre1day"), death)
```


### SOFA alone

```{r}
myTreeFunc(var = "SOFA" ,Pre1day_total_score_num, death)
```


### Categorical SOFA alone

```{r}
myTreeFunc(var = "SOFA categories" ,FourGrp_Pre1day_total_score_num, death)
```


### Pulmonary sofa alone

```{r}
myTreeFunc(var = "Pulmonary" ,Pre1day_respiratory_sofa_points_num, death)
```


### Renal sofa alone

```{r}
myTreeFunc(var = "Renal" ,Pre1day_bilirubin_sofa_points_num, death)
```

### Heme sofa alone

```{r}
myTreeFunc(var = "Heme" ,Pre1day_platelet_sofa_points_num, death)
```

### Neuro sofa alone

```{r}
myTreeFunc(var = "Neuro" ,Pre1day_gcs_sofa_points_num, death)
```

### Liver sofa alone

```{r}
myTreeFunc(var = "Liver" ,Pre1day_creatinine_sofa_points_num, death)
```

### Cardio sofa alone

```{r}
myTreeFunc(var = "Cardio" ,Pre1day_cardiovascular_sofa_points_num, death)
```





### SOFA + all components + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "SOFA & all components + covariates" ,contains("Pre1day"), AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```


### SOFA + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "SOFA + covariates" ,Pre1day_total_score_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```


### Pulmonary sofa + age + gender + obesity +diabetes + hypertension 

```{r}
myTreeFunc(var = "Pulmonary sofa + covariates" ,Pre1day_respiratory_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```


### Renal sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Renal sofa + covariates" ,Pre1day_bilirubin_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Heme sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Heme sofa + covariates" ,Pre1day_platelet_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Neuro sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Neuro sofa + covariates" ,Pre1day_gcs_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Liver sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Liver sofa + covariates" ,Pre1day_creatinine_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Cardio sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Cardio sofa + covariates" ,Pre1day_cardiovascular_sofa_points_num,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```



### age

```{r}
myTreeFunc(var = "age" ,AGE, death)
```

### age + SOFA

```{r}
myTreeFunc(var = "age + SOFA" ,AGE, Pre1day_total_score_num, death)
```

### age + Categorical SOFA

```{r}
myTreeFunc(var = "age + Categorical SOFA" ,FourGrp_Pre1day_total_score_num, AGE, death)
```

###  age + elixhauser score

```{r}
myTreeFunc(var = "age + elixhauser score" ,AGE, ElixhauserScore, death)
```

###  age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "covariates" ,AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### SOFA + elixhauser score

```{r}
myTreeFunc(var = "SOFA + elixhauser score" ,Pre1day_total_score_num, ElixhauserScore,death)
```

### Categories SOFA + elixhauser score

```{r}
myTreeFunc(var = "Categories SOFA + elixhauser score" ,FourGrp_Pre1day_total_score_num, ElixhauserScore,death)
```

### SOFA + age + elixhauser score

```{r}
myTreeFunc(var = "SOFA + age + elixhauser score" ,Pre1day_total_score_num, AGE, ElixhauserScore,death)
```

### SOFA + all components + age + elixhauser score

```{r}
myTreeFunc(var = "SOFA + all components + age + elixhauser score" ,contains("Pre1day"), AGE, ElixhauserScore,death)
```


### Pulmonary sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Pulmonary sofa + age + elixhauser score" ,Pre1day_respiratory_sofa_points_num, AGE, ElixhauserScore,death)
```


### Renal sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Renal sofa + age + elixhauser score" ,Pre1day_bilirubin_sofa_points_num, AGE, ElixhauserScore,death)
```

### Heme sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Heme sofa + age + elixhauser score" ,Pre1day_platelet_sofa_points_num, AGE, ElixhauserScore,death)
```

### Neuro sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Neuro sofa + age + elixhauser score" ,Pre1day_gcs_sofa_points_num, AGE, ElixhauserScore,death)
```

### Liver sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Liver sofa + age + elixhauser score" ,Pre1day_creatinine_sofa_points_num, AGE, ElixhauserScore,death)
```

### Cardio sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Cardio sofa + age + elixhauser score" ,Pre1day_cardiovascular_sofa_points_num, AGE, ElixhauserScore,death)
```




## GLM model

```{r}
myGLMFunc <- function(var,...){
  
glm0 <- glm(death~ .,data=FirstMV.select%>%select(...),family=binomial())
glmout <- broom::tidy(glm0, exponentiate = TRUE, conf.int=TRUE )%>%
   mutate_if(is.numeric, round, 2)%>%
  filter(term!="(Intercept)")%>%
  mutate(OR = paste0(round(estimate, 2), " (",round(conf.low,2),", ",round(conf.high,3),")"))

# Evaluate model performance On training set
prob.discharge <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
roc.train <-roc(FirstMVdata.train$death, prob.discharge)
ci.train <- ci.auc(roc.train)
auc.train <- paste0(round(as.numeric(roc.train$auc),2), " (",round(as.numeric(ci.train)[1],2),",",round(as.numeric(ci.train)[3],2),")")

# Evaluate model performance On test set
prob.discharge <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)
roc.test <- roc(FirstMVdata.test$death, prob.discharge)
ci.test<- ci.auc(roc.test)
auc.test <- paste0(round(as.numeric(roc.test$auc),2), " (",round(as.numeric(ci.test)[1],2),",",round(as.numeric(ci.test)[3],2),")")

out <- data.frame(Variable = var,AUC.train = auc.train, AUC.test = auc.test, term = glmout$term, OR = glmout$OR)
return(kable(out))
}
```

### SOFA alone

```{r}
myGLMFunc(var = "SOFA", Pre1day_total_score_num, death)
```

### Categorical SOFA alone

```{r}
myGLMFunc(var = "Categorical SOFA", FourGrp_Pre1day_total_score_num, death)
```

### Pulmonary +Renal + Heme + Neuro + Liver + Cardio 

```{r}
myGLMFunc(var = "All SOFA components", contains("Pre1day"),-Pre1day_total_score_num, -FourGrp_Pre1day_total_score_num,death)
```

### SOFA + age + gender + obesity +diabetes + hypertension

```{r}
myGLMFunc(var = "SOFA + covariates", Pre1day_total_score_num, AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### Pulmonary +Renal + Heme + Neuro + Liver + Cardio + age + gender + obesity +diabetes + hypertension

```{r}
myGLMFunc(var = "All SOFA components + covariates", contains("Pre1day"),-Pre1day_total_score_num,-FourGrp_Pre1day_total_score_num,
          AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### SOFA + age + elixhauser score

```{r}
myGLMFunc(var = "SOFA + age + elixhauser score", Pre1day_total_score_num, AGE, ElixhauserScore, death)
```

### Pulmonary +Renal + Heme + Neuro + Liver + Cardio + age + elixhauser score

```{r}
myGLMFunc(var = "SOFA components + age + elixhauser score", contains("Pre1day"),-Pre1day_total_score_num,-FourGrp_Pre1day_total_score_num,
          AGE, ElixhauserScore,death)
```

### age

```{r}
myGLMFunc(var = "age", AGE, death)
```

### age + SOFA

```{r}
myGLMFunc(var = "age + SOFA", AGE, Pre1day_total_score_num, death)
```

### age + Categories SOFA

```{r}
myGLMFunc(var = "age + Categories SOFA", AGE, FourGrp_Pre1day_total_score_num, death)
```

### age + Elixhauser score

```{r}
myGLMFunc(var = "age + Elixhauser score",  AGE, ElixhauserScore, death)
```

### SOFA + Elixhauser score

```{r}
myGLMFunc(var = "SOFA + Elixhauser score", Pre1day_total_score_num, ElixhauserScore,death)
```

### Categories SOFA + Elixhauser score

```{r}
myGLMFunc(var = "Categories SOFA + Elixhauser score", FourGrp_Pre1day_total_score_num, ElixhauserScore, death)
```



## Calibration plot

Produced calibration plots for logistic regression models using Age, SOFA, SOFA categories, SOFA + Age, or SOFA + Covariates as the regressors. 

```{r}
FirstMVdata.train<- FirstMVdata.train%>%
  mutate(Outcome = case_when(
    death =="deceased" ~ 1,
    death == "discharged" ~ 0
  ))

FirstMV.select<- FirstMV.select%>%
  mutate(Outcome = case_when(
    death =="deceased" ~ 1,
    death == "discharged" ~ 0
  ))

FirstMVdata.test<- FirstMVdata.test%>%
  mutate(Outcome = case_when(
    death =="deceased" ~ 1,
    death == "discharged" ~ 0
  ))

### SOFA alone
glm0 <- glm(Outcome~ .,data=FirstMV.select%>%select(Pre1day_total_score_num, Outcome),family=binomial())

prob.discharge.train <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
prob.discharge.test <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)

cb.SOFA <- givitiCalibrationBelt(o = FirstMVdata.train$Outcome, e = prob.discharge.train, devel = "internal", confLevels =0.95)
cb.SOFA.test <- givitiCalibrationBelt(o = FirstMVdata.test$Outcome, e = prob.discharge.test, devel = "external", confLevels =0.95)

roc.train.sofa <- roc(FirstMVdata.train$Outcome, prob.discharge.train, percent=TRUE)
roc.test.sofa <- roc(FirstMVdata.test$Outcome, prob.discharge.test, percent=TRUE)


### SOFA categories
glm0 <- glm(Outcome~ .,data=FirstMV.select%>%select(FourGrp_Pre1day_total_score_num, Outcome)%>%
              mutate(FourGrp_Pre1day_total_score_num= as.factor(FourGrp_Pre1day_total_score_num)),family=binomial())

prob.discharge.train <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
prob.discharge.test <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)

cb.SOFACate <- givitiCalibrationBelt(o = FirstMVdata.train$Outcome, e = prob.discharge.train, devel = "internal", confLevels =0.95)
cb.SOFACate.test <- givitiCalibrationBelt(o = FirstMVdata.test$Outcome, e = prob.discharge.test, devel = "external", confLevels =0.95)

roc.train.sofaCat <- roc(FirstMVdata.train$Outcome, prob.discharge.train, percent=TRUE)
roc.test.sofaCat <- roc(FirstMVdata.test$Outcome, prob.discharge.test, percent=TRUE)


### SOFA +Age
glm0 <- glm(Outcome~ .,data=FirstMV.select%>%select(Pre1day_total_score_num, AGE, Outcome),family=binomial())

prob.discharge.train <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
prob.discharge.test <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)

cb.SOFA.Age <- givitiCalibrationBelt(o = FirstMVdata.train$Outcome, e = prob.discharge.train, devel = "internal", confLevels =0.95)
cb.SOFA.Age.test <- givitiCalibrationBelt(o = FirstMVdata.test$Outcome, e = prob.discharge.test, devel = "external", confLevels =0.95)

roc.train.sofaAge <- roc(FirstMVdata.train$Outcome, prob.discharge.train, percent=TRUE)
roc.test.sofaAge <- roc(FirstMVdata.test$Outcome, prob.discharge.test, percent=TRUE)


### SOFA + Covariates (age + gender + obesity +diabetes + hypertension)
glm0 <- glm(Outcome~ .,data=FirstMV.select%>%select(Pre1day_total_score_num, AGE, GENDER, Obesity, Diabetes, Hyptertension,Outcome),family=binomial())

prob.discharge.train <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
prob.discharge.test <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)

cb.SOFA.Cov <- givitiCalibrationBelt(o = FirstMVdata.train$Outcome, e = prob.discharge.train, devel = "internal", confLevels =0.95)
cb.SOFA.Cov.test <- givitiCalibrationBelt(o = FirstMVdata.test$Outcome, e = prob.discharge.test, devel = "external", confLevels =0.95)


### Age
glm0 <- glm(Outcome~ .,data=FirstMV.select%>%select(AGE, Outcome),family=binomial())

prob.discharge.train <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
prob.discharge.test <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)

cb.Age <- givitiCalibrationBelt(o = FirstMVdata.train$Outcome, e = prob.discharge.train, devel = "internal", confLevels =0.95)
cb.Age.test <- givitiCalibrationBelt(o = FirstMVdata.test$Outcome, e = prob.discharge.test, devel = "external", confLevels =0.95)

roc.train.age <- roc(FirstMVdata.train$Outcome, prob.discharge.train, percent=TRUE)
roc.test.age <- roc(FirstMVdata.test$Outcome, prob.discharge.test, percent=TRUE)


pdf("CalibrationPlot_OnTrainTest_Main_removedImputedZero.pdf", height = 8, width = 8)

plot(cb.SOFA, main = "SOFA (continuous), training set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")
plot(cb.SOFA.test, main = "SOFA (continuous), validation set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")

plot(cb.SOFACate, main = "SOFA categories, training set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")
plot(cb.SOFACate.test, main = "SOFA categories, validation set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")


plot(cb.SOFA.Age, main = "SOFA + Age, training set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")
plot(cb.SOFA.Age.test, main = "SOFA + Age, validation set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")


plot(cb.SOFA.Cov, main = "SOFA + Covariates (age + gender + obesity +diabetes + hypertension), training set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")
plot(cb.SOFA.Cov.test, main = "SOFA + Covariates (age + gender + obesity +diabetes + hypertension), validation set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")

plot(cb.Age, main = "Age, training set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")
plot(cb.Age.test, main = "Age, validation set",
                          xlab = "Predicted probability",
                          ylab = "Observed mortality")

dev.off()
```


