---
title: "EHR SOFA --- Table 2, Supplemental Table 3a"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: "H:\\word rMarkDown template.docx"
---

```{r global_options, include=FALSE}
library(partykit)
library(dplyr)
library(pROC)
library(ROCR)
library(caret)
library(RColorBrewer)
library(flextable)
library(rmarkdown)
library(knitr)
library(readr)

library(givitiR)
library(gridExtra)
library(ggplot2)

library(gtsummary)
library(flextable)
library(officer)

load(here::here("FirstMV_20JUL2021_final.RData"))

opts_chunk$set(fig.width=10, fig.height=8,echo=FALSE, warning=FALSE, message=FALSE, comment = NA)
```

```{r package_options, include=FALSE}
opts_knit$set(progress = TRUE, verbose = TRUE)
```


## Conditional Tree

```{r}

FirstMV.select <- FirstMV.select%>%mutate_if(is.character, as.factor)
FirstMVdata.train <- FirstMVdata.train%>%mutate_if(is.character, as.factor)
FirstMVdata.test <- FirstMVdata.test%>%mutate_if(is.character, as.factor)


myTreeFunc <- function(var,...){
  
set.seed(123)
temp0 <- (ctree(death~.,data=FirstMV.select%>%select(...),minbucket=1000))

plot(temp0, tp_args = list(fill =c("darkgreen","red")))

# Evaluate model performance On training set
prob.discharge <- predict(temp0, newdata=FirstMVdata.train, type="prob",simplify=T)[,2]
roc.train <- roc(FirstMVdata.train$death, prob.discharge)
ci.train <- ci.auc(roc.train)
auc.train <- paste0(round(as.numeric(roc.train$auc),3), " (",round(as.numeric(ci.train)[1],3),",",round(as.numeric(ci.train)[3],3),")")

# evaluate model performance On test set
prob.discharge <- predict(temp0, newdata=FirstMVdata.test, type="prob",simplify=T)[,2]
roc.test <- roc(FirstMVdata.test$death, prob.discharge)
ci.test<- ci.auc(roc.test)
auc.test <- paste0(round(as.numeric(roc.test$auc),3), " (",round(as.numeric(ci.test)[1],3),",",round(as.numeric(ci.test)[3],3),")")

# confusion matrix 
# predict.discharge <- predict(temp0, newdata=FirstMVdata.test, type="response",simplify=T)
# CM <-confusionMatrix(as.factor(FirstMVdata.test$death),predict.discharge)
# CM$table
out <- data.frame(Variable = var,AUC.train = auc.train, AUC.test = auc.test)
return(kable(out))
}
```



### SOFA + all components

```{r}
myTreeFunc(var = "SOFA + all components" ,contains("Pre1day"), death)
```


### SOFA alone

```{r}
myTreeFunc(var = "SOFA" ,Pre1day_total_score_num_imp, death)
```


### Categorical SOFA alone

```{r}
myTreeFunc(var = "SOFA categories" ,FourGrp_Pre1day_total_score_num_imp, death)
```


### Pulmonary sofa alone

```{r}
myTreeFunc(var = "Pulmonary" ,Pre1day_respiratory_sofa_points_num_imp, death)
```


### Renal sofa alone

```{r}
myTreeFunc(var = "Renal" ,Pre1day_bilirubin_sofa_points_num_imp, death)
```

### Heme sofa alone

```{r}
myTreeFunc(var = "Heme" ,Pre1day_platelet_sofa_points_num_imp, death)
```

### Neuro sofa alone

```{r}
myTreeFunc(var = "Neuro" ,Pre1day_gcs_sofa_points_num_imp, death)
```

### Liver sofa alone

```{r}
myTreeFunc(var = "Liver" ,Pre1day_creatinine_sofa_points_num_imp, death)
```

### Cardio sofa alone

```{r}
myTreeFunc(var = "Cardio" ,Pre1day_cardiovascular_sofa_points_num_imp, death)
```





### SOFA + all components + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "SOFA & all components + covariates" ,contains("Pre1day"), AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```


### SOFA + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "SOFA + covariates" ,Pre1day_total_score_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```


### Pulmonary sofa + age + gender + obesity +diabetes + hypertension 

```{r}
myTreeFunc(var = "Pulmonary sofa + covariates" ,Pre1day_respiratory_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```


### Renal sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Renal sofa + covariates" ,Pre1day_bilirubin_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Heme sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Heme sofa + covariates" ,Pre1day_platelet_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Neuro sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Neuro sofa + covariates" ,Pre1day_gcs_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Liver sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Liver sofa + covariates" ,Pre1day_creatinine_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```

### Cardio sofa + age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "Cardio sofa + covariates" ,Pre1day_cardiovascular_sofa_points_num_imp,AGE, GENDER, Obesity, Diabetes, Hyptertension, death)
```



### age

```{r}
myTreeFunc(var = "age" ,AGE, death)
```

### age + SOFA

```{r}
myTreeFunc(var = "age + SOFA" ,AGE, Pre1day_total_score_num_imp, death)
```

### age + Categorical SOFA

```{r}
myTreeFunc(var = "age + Categorical SOFA" ,FourGrp_Pre1day_total_score_num_imp, AGE, death)
```

###  age + elixhauser score

```{r}
myTreeFunc(var = "age + elixhauser score" ,AGE, ElixhauserScore, death)
```

###  age + gender + obesity +diabetes + hypertension

```{r}
myTreeFunc(var = "covariates" ,AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### SOFA + elixhauser score

```{r}
myTreeFunc(var = "SOFA + elixhauser score" ,Pre1day_total_score_num_imp, ElixhauserScore,death)
```

### Categories SOFA + elixhauser score

```{r}
myTreeFunc(var = "Categories SOFA + elixhauser score" ,FourGrp_Pre1day_total_score_num_imp, ElixhauserScore,death)
```

### SOFA + age + elixhauser score

```{r}
myTreeFunc(var = "SOFA + age + elixhauser score" ,Pre1day_total_score_num_imp, AGE, ElixhauserScore,death)
```

### SOFA + all components + age + elixhauser score

```{r}
myTreeFunc(var = "SOFA + all components + age + elixhauser score" ,contains("Pre1day"), AGE, ElixhauserScore,death)
```


### Pulmonary sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Pulmonary sofa + age + elixhauser score" ,Pre1day_respiratory_sofa_points_num_imp, AGE, ElixhauserScore,death)
```


### Renal sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Renal sofa + age + elixhauser score" ,Pre1day_bilirubin_sofa_points_num_imp, AGE, ElixhauserScore,death)
```

### Heme sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Heme sofa + age + elixhauser score" ,Pre1day_platelet_sofa_points_num_imp, AGE, ElixhauserScore,death)
```

### Neuro sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Neuro sofa + age + elixhauser score" ,Pre1day_gcs_sofa_points_num_imp, AGE, ElixhauserScore,death)
```

### Liver sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Liver sofa + age + elixhauser score" ,Pre1day_creatinine_sofa_points_num_imp, AGE, ElixhauserScore,death)
```

### Cardio sofa + age + elixhauser score

```{r}
myTreeFunc(var = "Cardio sofa + age + elixhauser score" ,Pre1day_cardiovascular_sofa_points_num_imp, AGE, ElixhauserScore,death)
```




## GLM model

```{r}
myGLMFunc <- function(var,...){
  
glm0 <- glm(death~ .,data=FirstMV.select%>%select(...),family=binomial())
glmout <- broom::tidy(glm0, exponentiate = TRUE, conf.int=TRUE )%>%
   mutate_if(is.numeric, round, 2)%>%
  filter(term!="(Intercept)")%>%
  mutate(OR = paste0(round(estimate, 2), " (",round(conf.low,2),", ",round(conf.high,3),")"))

# Evaluate model performance On training set
prob.discharge <- predict(glm0,  newdata=FirstMVdata.train, type="response",simplify=T)
roc.train <-roc(FirstMVdata.train$death, prob.discharge)
ci.train <- ci.auc(roc.train)
auc.train <- paste0(round(as.numeric(roc.train$auc),2), " (",round(as.numeric(ci.train)[1],2),",",round(as.numeric(ci.train)[3],2),")")

# Evaluate model performance On test set
prob.discharge <- predict(glm0, newdata=FirstMVdata.test, type="response",simplify=T)
roc.test <- roc(FirstMVdata.test$death, prob.discharge)
ci.test<- ci.auc(roc.test)
auc.test <- paste0(round(as.numeric(roc.test$auc),2), " (",round(as.numeric(ci.test)[1],2),",",round(as.numeric(ci.test)[3],2),")")

out <- data.frame(Variable = var,AUC.train = auc.train, AUC.test = auc.test, term = glmout$term, OR = glmout$OR)
return(kable(out))
}
```

### SOFA alone

```{r}
myGLMFunc(var = "SOFA", Pre1day_total_score_num_imp, death)
```

### Categorical SOFA alone

```{r}
myGLMFunc(var = "Categorical SOFA", FourGrp_Pre1day_total_score_num_imp, death)
```

### Pulmonary 

```{r}
myGLMFunc(var = "Pulmonary sofa", Pre1day_respiratory_sofa_points_num_imp,death)
```

### Renal 

```{r}
myGLMFunc(var = "Renal sofa",Pre1day_bilirubin_sofa_points_num_imp,death)
```

###  Heme

```{r}
myGLMFunc(var = "Hemo sofa", Pre1day_platelet_sofa_points_num_imp,death)
```

### Neuro

```{r}
myGLMFunc(var = "Neuro sofa", Pre1day_gcs_sofa_points_num_imp,death)
```

### Liver

```{r}
myGLMFunc(var = "Liver sofa", Pre1day_creatinine_sofa_points_num_imp,death)
```

### Cardio 

```{r}
myGLMFunc(var = "Cardio sofa", Pre1day_cardiovascular_sofa_points_num_imp,death)
```







### Pulmonary +Renal + Heme + Neuro + Liver + Cardio 

```{r}
myGLMFunc(var = "All SOFA components", contains("Pre1day"),-Pre1day_total_score_num_imp, -FourGrp_Pre1day_total_score_num_imp,death)
```

### SOFA + age + gender + obesity +diabetes + hypertension

```{r}
myGLMFunc(var = "SOFA + covariates", Pre1day_total_score_num_imp, AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### Pulmonary +Renal + Heme + Neuro + Liver + Cardio + age + gender + obesity +diabetes + hypertension

```{r}
myGLMFunc(var = "All SOFA components + covariates", contains("Pre1day"),-Pre1day_total_score_num_imp,-FourGrp_Pre1day_total_score_num_imp,
          AGE, GENDER, Obesity, Diabetes, Hyptertension,death)
```

### SOFA + age + elixhauser score

```{r}
myGLMFunc(var = "SOFA + age + elixhauser score", Pre1day_total_score_num_imp, AGE, ElixhauserScore, death)
```

### Pulmonary +Renal + Heme + Neuro + Liver + Cardio + age + elixhauser score

```{r}
myGLMFunc(var = "SOFA components + age + elixhauser score", contains("Pre1day"),-Pre1day_total_score_num_imp,-FourGrp_Pre1day_total_score_num_imp,
          AGE, ElixhauserScore,death)
```

### age

```{r}
myGLMFunc(var = "age", AGE, death)
```

### age + SOFA

```{r}
myGLMFunc(var = "age + SOFA", AGE, Pre1day_total_score_num_imp, death)
```

### age + Categories SOFA

```{r}
myGLMFunc(var = "age + Categories SOFA", AGE, FourGrp_Pre1day_total_score_num_imp, death)
```

### age + Elixhauser score

```{r}
myGLMFunc(var = "age + Elixhauser score",  AGE, ElixhauserScore, death)
```

### SOFA + Elixhauser score

```{r}
myGLMFunc(var = "SOFA + Elixhauser score", Pre1day_total_score_num_imp, ElixhauserScore,death)
```

### Categories SOFA + Elixhauser score

```{r}
myGLMFunc(var = "Categories SOFA + Elixhauser score", FourGrp_Pre1day_total_score_num_imp, ElixhauserScore, death)
```







